### 2020-10-01 analysis/paper/paper.tex
segudno borrador

%% Version 5.0, 2 January 2020
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% TemplateV5.tex --  LaTeX-based template for submissions to the
% American Meteorological Society
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PREAMBLE
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% Start with one of the following:
% DOUBLE-SPACED VERSION FOR SUBMISSION TO THE AMS
\documentclass[]{ametsocV5}

% TWO-COLUMN JOURNAL PAGE LAYOUT---FOR AUTHOR USE ONLY
% \documentclass[twocol]{ametsocV5}

% Enter packages here. If too many math alphabets are used,
% remove unnecessary packages or define hmmax and bmmax as necessary.

\newcommand{\hmmax}{0}
\newcommand{\bmmax}{0}
\usepackage{amsmath,amsfonts,amssymb,bm}
\usepackage{mathptmx}%{times}
\usepackage{newtxtext}
\usepackage{newtxmath}

\usepackage{gensymb}
\usepackage{subfig}
\usepackage[inline]{showlabels}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% To be entered by author:

%% May use \\ to break lines in title:

\title{Asssesment of zonal symmetric and asymmetric components of the Southern
Annular Mode using a novel approach}

\authors{Elio Campitelli
\correspondingauthor{Elio Campitelli,elio.campitelli@cima.fcen.uba.ar}
and Leandro B. Díaz
}

\affiliation{Universidad de Buenos Aires, Facultad de Ciencias Exactas y Naturales,
Departamento de Ciencias de la Atmósfera y los Océanos, Buenos Aires,
Argentina CONICET -- Universidad de Buenos Aires, Centro de
Investigaciones del Mar y la Atmósfera (CIMA), Buenos Aires, Argentina
CNRS -- IRD -- CONICET -- UBA, Instituto Franco‐Argentino para el
Estudio del Clima y sus Impactos (UMI 3351 IFAECI), Buenos Aires,
Argentina}

\extraauthor{Carolina Vera
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ABSTRACT
%
% Enter your abstract here
% Abstracts should not exceed 250 words in length!
%

\abstract{Enter the text of your abstract here. This is a sample American
Meteorological Society (AMS) \LaTeX~template. This document provides
authors with instructions on the use of the AMS \LaTeX~template. Authors
should refer to the file amspaper.tex to review the actual \LaTeX~code
used to create this document. The template.tex file should be modified
by authors for their own manuscript.}

\begin{document}

%% Necessary!
\maketitle

\bibliographystyle{ametsoc2014}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% SIGNIFICANCE STATEMENT/CAPSULE SUMMARY
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% If you are including an optional significance statement for a journal article or a required capsule summary for BAMS
% (see www.ametsoc.org/ams/index.cfm/publications/authors/journal-and-bams-authors/formatting-and-manuscript-components for details),
% please apply the necessary command as shown below:
%
\statement
This is significant becasue I wrote it.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% MAIN BODY OF PAPER
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%

\section{Introduction}

The Southern Annular Mode (SAM) is the main mode of variability in the
Southern Hemisphere extratropical circulation \citep{rogers1982} in
daily, monthly, and decadal timescales \citep{baldwin2001a, fogt2006}
and exerts an important influence in weather conditions such as
temperature and precipitation anomalies and sea ice concentration
\citep{fogt2020}. Its positive phase is traditionally described as
anomalously low pressures over Antarctica surrounded by a ring of
anomalous high pressures in middle-to-high latitudes.

Most authors describe the SAM as a zonally symmetric pattern, a fact
that is reflected not only in its name, but also in the various methods
used to characterise it. Of the several different indices presented in
the literature, many of them are based on zonal means of sea level
pressure or geopotential height \citep{ho2012}. \citet{gong1999} defined
the SAM index as the zonal mean sea level pressure difference between
40\degree S and 60\degree S, which is also the definition used by the
station-based index in \citet{marshall2003}. \citet{baldwin2009}
proposed defining the Northern and Annular modes as the leading EOF of
the zonally averaged geopotential height at each level.

Even though these indices are based on zonal averages, the spatial the
spatial structure of the SAM computed from them contains noticeable
deviations from zonal symmetry, particularly in the Pacific Ocean
region. These zonal asymmetries have not been widely studied, but
previous work suggest that they strongly modulate the regional impacts
of the SAM, going as far as reversing its relationship between
precipitation in South America \citep{silvestri2009}. At the very least,
the fact that the SAM is not entirely zonally symmetric hinders our
ability to reconstruct its historical variability prior to the
availability of dense observations in the Southern Hemisphere
\citep{jones2009}.

At least some of the variability associated with the zonal asymmetries
of the SAM is probably forced by the tropics. In particular, ENSO-like
variability affects the Southern Hemisphere extratopics through the
Pacific-South American Pattern \citep{mo1987, kidson1988, karoly1989},
whose wave train protects strongly onto the zonal anomalies
corresponding to the SAM in the Pacific sector. And although the
relationship between ENSO and SAM is far from simple, tropical
influences on the SAM have been observed
\citep{fan2007, fogt2011, clem2013}. In particular, \citet{fan2007}
computed SAM indices of the Western and the Eastern Hemisphere
separately and found that they were much more correlated if the (linear)
signal of the ENSO was removed. While this relates to temporally
coherent variability of the two hemisphere and not necessarily to zonal
asymmetries in the associated spatial patterns, it is nonetheless
consistent of the ENSO playing a crucial role in zonal asymmetries of
the SAM.

Positive trends in SAM index have been documented by various researchers
using different indices mostly on boreal Summer and Autumn \citep[ and
references therein]{fogt2020}. These trends are thought of driven
primarily by stratospheric ozone depletion and understood in the context
of zonal mean variables. However, it's not clear how or if the
asymmetric component responds to this forcing, or whether its
variability could be masking influencing the observed trends
independently.

Similarly unclear are the specific impacts of the zonally asymmetric
component of the SAM. Positive phase of the SAM is associated with
generally colder temperatures over Antarctica and warmer temperatures at
higher latitudes \citep{jones2019} (and vice versa for negative SAM),
but there are significant deviations from this zonal mean response,
notably in the Antarctic Peninsula and the South Atlantic
\citep{fogt2012}. The SAM signal in precipitation behaves similarly,
although with even greater deviation from zonal symmetry
\citep{lim2016}. The importance of zonal asymmetries of the SAM in these
impacts have been studied in certain regions. For example, the
SAM-precipitation relationship in Southeastern South America and
Southern Brazil can be explained by the PSA-like zonally asymmetric
circulation associated with the SAM \citep{silvestri2009, rosso2018}.
\citet{fan2007} also found that precipitation in East Asia was impacted
by the variability of only the Western Hemisphere part of the SAM.

We are not aware of any previous work which quantifies the temporal
variability of the asymmetric component of the SAM with the exception of
\citet{fogt2012}. However, their methods based on composites of positive
and negative SAM events leads to some issues, such as spatial patterns
derived from as little as 4 cases and from imbalanced periods (for
example, 5 of the 7 cases in their DJF SAM+ composite are from later
than 1988, whereas all of the 8 years in their DJF SAM- composite are
from earlier than 1988). This is particularly important due to the
inhomogeneities in reanalysis products prior to the satellite era and
the possible change in the asymmetric structure of the SAM
\citep{silvestri2009}.

Moreover, \citet{fogt2012} studied the zonal asymmetric component of the
SAM only in sea level pressure. Zonal asymmetries in the SAM spatial
pattern are fairly barotropic throughout the troposphere, but they
change dramatically in the stratosphere \citep{baldwin2009}.

Our objective is, then, to systematically characterise the zonally
asymmetric and symmetric components of the SAM variability. For each
level, we construct two indices which aim to capture exclusively the
variability of the symmetric and asymmetric component respectively. We
assess their vertical structure and coherence, temporal variability and
trends. We then study the spatial patterns described by the variability
exclusive to each index. Finally, we investigate their relationship with
temperature and precipitation anomalies.

In the Section \ref{sec:methods} we describe the methods. In Section
\ref{sec:results}.\ref{sec:temporal} we describe the temporal
variability and vertical coherence of the indices. In Section
\ref{sec:results}.\ref{sec:spatial}, we analyse the spatial patterns of
geopotential height associated with them. In Section
\ref{sec:results}.\ref{sec:impacts}, we study their relationship with
surface-level temperature and precipitation.

\section{Methods}

\label{sec:methods}

\subsubsection{Data}

To describe the Southern Annular Mode and its variability we used
monthly geopotential height at 2.5\degree~longitude by
2.5\degree~latitude of horizontal resolution and 37 vertical isobaric
levels from ERA5 \citep{hersbach2020} for the period 1979 to 2018. We
restrict our analysis to the post-satellite era to avoid any confounding
factors arising from the introduction of satellite observations.

To describe the relationship between the SAM indices and temperature and
precipitation, we use temperature data from NOAA's Merged Land Ocean
Global Surface Temperature Analysis V4.0.1 \citep{vose2012, smith2008},
which blends land and ocean temperature analysis into a monthly global
grid 5\degree~longitude by 5\degree~latitude, and monthly rainfall
0.5\degree~longitude by 0.5\degree~latitude data from the Global
Precipitation Climatology Centre \citep{schneider2015}.

\subsubsection{Definition of indices}

Traditionally the Souther Annular Mode (SAM) is defined as de leading
Empirical Orthogonal Mode (EOF) of sea level pressure or geopotential
height at lower levels \citep{ho2012}. Following \citet{baldwin2001}, we
extend that definition vertically and use the term SAM to refer to the
the leading EOF of the monthly anomalies of geopotential field south of
20\degree S at each level. We performed EOFs by computing the Singular
Value Decomposition of the data matrix consisting in 481 rows and 4176
columns (144 points of longitude and 29 points of latitude). We weighted
the values by the square root of the cosine of latitude to account for
the non-equal area of each gridpoint \citep{chung1999}.

To separate the zonally symmetric and asymmetric components of the SAM,
we computed the zonal mean and anomalies of the full SAM spatial
pattern, as shown in Figure \ref{fig:method} for 700 hPa. The full
spatial signal (\(\mathrm{EOF_1}(\lambda, \phi)\)) is the sum of the
zonally asymmetric (\(\mathrm{EOF_1^*}(\lambda, \phi)\)) and symmetric
(\([\mathrm{EOF_1}](\lambda, \phi)\)) components. We then compute the
``Full SAM'', ``Asymmetric SAM'' and ``Symmetric SAM'' indices as the
regression coefficients of the regression of each monthly geopotential
field on the respective patterns (weighting by the cosine of latitude).
The three indices are normalized by dividing them by the standard
deviation of the ``Full'' index at each level. As a result, the
magnitudes between indices are comparable. However, only ``Full'' index
will have unit standard deviation per definition. From the regression,
we also use the explained variance of each pattern as an indicator of
the degree of zonally symmetry or asymmetry of each monthly field.

\begin{figure*}
\includegraphics{method-1} \caption[Spatial patterns of the first EOF of 700 hPa geopotential height for 1979–2018 period]{Spatial patterns of the first EOF of 700 hPa geopotential height for 1979–2018 period. (a) Full field, (b) zonally asymmetric component and (c) zonally symmetric component. Arbitrary units.}\label{fig:method}
\end{figure*}

Our method assumes linearity in the asymmetric component of the SAM.
That is, we assume that zonal symmetries associated with positive SAM
are almost opposite and of the same magnitude to the ones associated
with negative SAM. \citet{fogt2012}'s composites (their Figure 4)
suggest that this might not be entirely valid, although we argue that
much of that apparent non-linearity is due to the heterogenous nature of
the selected years for constructing the composites. Using our data (from
1979 to 2018), seasonal composites of zonal anomalies of 700 hPa
geopotential height for SAM+ (Full SAM index greater than 1 standard
deviation) and SAM- (less than minus 1 standard deviation) show
relatively high pattern correlations for all seasons and are visually
very linear (Figure A9). Therefore, we believe that our method is a
reasonable approximation of the phenomenon.

By computing a single EOF pattern using data for all months we are
assuming that the zonal anomalies of the SAM are the same in all seasons
--December to February (DJF), March to May (MAM), June to August (JJA)
and September to November (SON). Geopotential zonal anomalies computed
by projecting the first EOF \emph{of each season} are very similar to
each other (Figure A10) and show pattern correlations between 0.65 (DJF
with JJA) and 0.9 (between MAM and SON). Based on this, we believe that
our initial assumption is not unreasonable.

Finally, we assume that the zonally asymmetric pattern is stationary in
time. \citet{silvestri2009} suggest that this might not be the case
between 1958 and 2004 but the period we analyse is much shorter
(1979-2018) so it's unlikely that we could observe significant changes.
Moreover, zonal asymmetry of the spatial patterns for the two halves of
the period (1979 to 1998 and 1999 to 2018) show no systematic change
(Figure A11).

\subsubsection{Regressions}

We perform linear regressions to quantify the association between the
SAM indices and other variables. Since the Asymmetric and Symmetric SAM
indices are significantly correlated with each other, to capture the
variability explained uniquely by each index we use one multiple linear
regression instead of two simple linear regressions. To obtain the
linear coefficients of a variable \(X\) (geopotential, temperature,
precipitation, etc\ldots{}) with the Asymmetric SAM (\(SAM_a\)) and
Symmetric SAM (\(SAM_s\)) we fit the equation

\[
X(\lambda, \phi, t) = \alpha(\lambda, \phi) SAM_a + \beta(\lambda, \phi) SAM_s + X_0(\lambda, \phi) +  \epsilon(\lambda, \phi, t)
\]

where \(\lambda\) and \(\phi\) are the longitude and latitude, \(t\) is
the time, \(\alpha\) and \(\beta\) are the linear regression
coefficients, \(X_0\) and \(\epsilon\) are the constant and error terms.
From this equation, \(\alpha\) represents the (linear) association of
\(X\) with the variability of the Asymmetric SAM that is not explained
by the variability of the Symmetric SAM; i.e.~it is proportional to the
partial correlation of \(X\) and the Asymmetric SAM, controlling for the
effect of the Symmetric SAM and vice versa for \(\beta\). When
performing a separate regression for each trimester (DJF, MAM, JJA,
SON), we average of the relevant variables seasonally for each year and
trimester before computing the regression.

At 2.5\degree by 2.5\degree resolution, a single regression field is
composed of thousands of regressions. In such case, using p-values to
test for significance leads to misleading results
\citep{walker1914, katz1991}. While there are multiple proposed
solutions in the literature, \citet{wilks2016} suggests that adjusting
p-values by controlling for the False Discovery Rate
\citep{benjamini1995} is a simple and effective method to ameliorate
this issue. Therefore, p-values showed in regression fields are all
adjusted following \citet{benjamini1995}.

We computed linear trends by Ordinary Least Squares and the 95\%
confidence interval assuming a t-distribution of the appropriate
residual degrees of freedom.

To compute the amplitude of the zonal waves we computed the Fourier
transform of the spatial field at each latitude circle.

\subsubsection{Computation procedures}

We performed all analysis in this paper using the R programming language
\citep{rcoreteam2020}, using the data.table package \citep{dowle2020}
and the metR package \citep{campitelli2020}. All graphics are made using
ggplot2 \citep{wickham2009}. We downloaded data from ranalysis using the
ecmwfr package \citep{hufkens2020} and indices of the ENSO with the rsoi
package \citep{albers2020}. The paper was rendered using knitr and
rmarkdown \citep{xie2015, allaire2019}.

A version-controlled repository of the code used to create this analysis
can be found at https://github.com/eliocamp/asymsam.

\section{Results}

\label{sec:results}

\subsection{Temporal evolution}

\label{sec:temporal}

\begin{figure*}
\includegraphics{asymsam-timeseries-1} \caption[Time series for the Asymmetric SAM and Symmetric SAM indices at (a) 50 hPa and (b) 700 hPa]{Time series for the Asymmetric SAM and Symmetric SAM indices at (a) 50 hPa and (b) 700 hPa. To the left, probability density estimate of each index computed using a gaussian kernel of optimal bindwidth according to \cite{sheather1991}. Series are standarised by the standard deviation of the Full SAM at each level.}\label{fig:asymsam-timeseries}
\end{figure*}

We first asses the temporal evolution of the Asymmetric SAM and
Symmetric SAM. Figure \ref{fig:asymsam-timeseries} shows the
corresponding time series for 700 hPa and 50 hPa and their corresponding
density estimates. We selected these two levels as representative of the
tropospheric and stratospheric variability respectively. As will be
shown later, both indices are highly coherent within each atmospheric
layer, therefore is reasonable to take one level as representative of
each layer.

Month-to-month variability is evident for both indices, with noisy
variations in the low frequency. At first glance the series can be
distinguished by their distributions. Compared to the tropospheric
indices, the stratospheric indices are much more long-tailed; that is,
extreme values (both negative and positive) abound. The Asymmetric SAM
series have both more variability in the higher frequencies than the
Symmetric SAM series.

The stratospheric Symmetric SAM varies strongly with a two-year period,
which can be seen by spectral analysis (Figure A3). This might suggests
a link between stratospheric SAM variability and the Quasi-Biennial
Oscillation \citep{baldwin2001b}. There is a local peak at 2 years in
the periodogram of the tropospheric Symmetric SAM also, although it's
not statistically significant. In the troposphere the most significant
peak of variability is found in the Asymmetric index at around 3.6
months.

From Figure \ref{fig:asymsam-timeseries} we can see that the Asymmetric
SAM and Symmetric SAM time series appear to be correlated. Moreover,
looking at the extremes in the stratosphere, the Symmetric SAM series
appears to lag the Asymmetric SAM series (see, for example, the positive
events on late 1987). We show these correlations, across all the levels
of the reanalysis for zero and -1 lag (Asymmetric SAM index leading the
Symmetric SAM index), in Figure \ref{fig:cor-lev}. Zero-lag correlations
between the Asymmetric SAM and Symmetric SAM series are relatively
constant throughout the troposphere, fluctuating between 0.39 and 0.45.
One-month-lag correlations are similarly constant but significantly
reduced to around 0.17. In the stratosphere, zero-lag correlations drop
to a minimum of 0.21 at 20 hPa and then it increases again monotonically
with height up to the uppermost level of the reanalysis (although
results near the top of the models are to be interpreted with care). At
the same time, one-month-lag correlations increase with height. As a
consequence, stratospheric Asymmetric SAM index tend to precede
corresponding Symmetric SAM index.

\begin{figure}
\includegraphics{cor-lev-1} \caption[Correlation between the Symmetric SAM and Asymmetric SAM index at each level for lag zero and lag -1 (Asymmetric SAM leads Symmetric SAM) for the 1979–2018 period]{Correlation between the Symmetric SAM and Asymmetric SAM index at each level for lag zero and lag -1 (Asymmetric SAM leads Symmetric SAM) for the 1979–2018 period.}\label{fig:cor-lev}
\end{figure}

\begin{figure*}
\includegraphics{cross-correlation-1} \caption[Cross correlation between levels of the (a) Full SAM, (b) Asymmetric SAM, and (c) Symmetric SAM for the 1979–2018 period]{Cross correlation between levels of the (a) Full SAM, (b) Asymmetric SAM, and (c) Symmetric SAM for the 1979–2018 period.}\label{fig:cross-correlation}
\end{figure*}

Figure \ref{fig:cross-correlation}a shows (zero-lag) cross-correlation
across levels for the Full, Symmetric and Asymmetric SAM indices. For
the Full SAM (panel a), high values below 100 hPa reflect the vertical
(zero-lag) coherency throughout the troposphere. Above 100 hPa
correlation between levels falls off more rapidly, indicating less
coherent (zero-lag) variability. Therefore, there is a non negligible
correlation between the troposphere and the lower-to-middle
stratosphere. Examining panels b and c, we see that the Asymmetric and
Symmetric SAM share the same high level of coherency in the troposphere
but they differ in their stratospheric behaviour. Stratospheric
coherency is stronger for the Asymmetric SAM than the Symmetric SAM. The
stratospheric Symmetric SAM seems to connect more strongly to the
troposphere than the Asymmetric SAM.

\begin{figure*}
\includegraphics{trends-1} \caption[Decadal linear trends at each level for annual (row 1) and seasonal values (rows 2 to 5) for the period 1979-2018 and for the (column a) Full SAM index, (column b) Asymmetric SAM index, and (column c) Symmetric SAM index]{Decadal linear trends at each level for annual (row 1) and seasonal values (rows 2 to 5) for the period 1979-2018 and for the (column a) Full SAM index, (column b) Asymmetric SAM index, and (column c) Symmetric SAM index. Shading indicates the 95\% confidence interval from a t-distribution.}\label{fig:trends}
\end{figure*}

The linear trends for each of the indices (Full SAM, Symmetric SAM and
Asymmetric SAM) were evaluated for the complete period 1979-2018 at each
level (Figure \ref{fig:trends}) for the whole year and separated by
trimesters. The Full SAM index presents a statistically significant
trend (panel a.1) that extends throughout the troposphere up to about 50
hPa and reaches its maximum value at 100 hPa. The seasonal trends (rest
of column a) indicate that positive trends are present in Autumn and
particularly in Summer, where the 100 hPa maximum is much more defined.
In Winter and Spring, we detect no statistically significant trend. This
is consistent with the results of previous studies, which find clear
positive trends in Summer, weaker in Autumn and no trends in the other
seasons \citep[e.g.][ and references therein]{fogt2020} using indices of
the SAM based on surface or near-surface circulation.

By separating the SAM signal in its asymmetric and symmetric parts, we
can not only see that these trends are almost entirely due to the
symmetric component (column b vs.~column c), but in some cases the
trends become more clear. In Summer, the Asymmetric SAM has a
statistically non significant negative trend in the middle troposphere
that obscures the trend in the Full SAM index; as a result, trends
computed using only the Symmetric component are more clear (compare the
shading region in panel a.2 and c.2). In Autumn, the Symmetric SAM
reveals a statistically significant positive trend in the stratosphere
that is not significant using the Full SAM index.

We stress that these are only linear trends during the whole period and
the absence of a statistically significant signal should not be taken as
evidence of no systematic change. In particular, going back to Figure
\ref{fig:asymsam-timeseries}, we can see an evident change in the
stratospheric Asymmetric component (red line in panel a) between the
90's, when we see a dominance of extreme negative values, and the 00's,
when we see the inverse. This change is restricted to the Winter months:
the linear trend for Winter starting in 1990 for the Asymmetric
component at 50hPa is \(0.37 \pm 0.22\).

Figure \ref{fig:r-squared-trend} shows decadal trends for the explained
variance of each index. There is no evidence of a significant trend in
the stratosphere. In the troposphere, there is a positive trend for the
Asymmetric SAM and not significant trend for the Symmetric SAM. This
suggest that the SAM has become more asymmetric in the period from 1979
to 2018. However, the change is slight, around 1\% increased explained
variance per decade.

\begin{figure}
\includegraphics{r-squared-trend-1} \caption[Decadal trends of the variance explained by the Asymmetric and Symmetric SAM at each level for the period 1979-2018]{Decadal trends of the variance explained by the Asymmetric and Symmetric SAM at each level for the period 1979-2018. Shading indicates the 95\% confidence interval.}\label{fig:r-squared-trend}
\end{figure}

\subsection{Spatial patterns}

\label{sec:spatial}

\begin{figure*}
\includegraphics{2d-regr-1} \caption[Regression of geopotential height (meters) at (row 1) 50 hPa and (row 2) 700 hPa with the (column a) Full SAM, (column b) Asymmetric SAM, and (column c) Symmetric SAM for the 1979–2018 period]{Regression of geopotential height (meters) at (row 1) 50 hPa and (row 2) 700 hPa with the (column a) Full SAM, (column b) Asymmetric SAM, and (column c) Symmetric SAM for the 1979–2018 period. The regression patterns for Asymmetric and Symmetric SAM are the result of one multiple regression using both indices. Points marked on panel b.2 are the location of the reference points used by \cite{raphael2004} for their Zonal Wave 3 index. }\label{fig:2d-regr}
\end{figure*}

To show if, and to what extent, the Asymmetric and Symmetric SAM indices
indeed capture the asymmetric and symmetric component of the SAM
respectively, we computed the spatial regression of geopotential height
anomalies on these indices and the Full SAM index for 700 hPa and 50 hPa
levels. Figure \ref{fig:2d-regr} shows these regressions. Regression
coefficients in column a are computed using the Full SAM. Regression
coefficients in columns b and c are computed using multiple regression
using the Asymmetric and Symmetric indices at the same time. Thus, they
are to be interpreted as the patterns associated with each index,
removing the variability (linearly) explained by the other index.

In the stratosphere, the spatial pattern associated with the Full SAM is
more clearly dominated by a zonally symmetric, monopolar structure
(panel a.1) which is, however, not perfectly centred in the South Pole.
The monopole obtained by the regression pattern for Symmetric SAM (panel
c.1) is much more symmetric and the shift from total symmetry is
captured by the regression pattern of the Asymmetric SAM as a wave-1
with maximum anomalies above the Belinghausen Sea on the Western
Hemisphere and and Davis Sea in the Eastern Hemisphere (panel b.1).

In the troposphere, panel a.2 shows the well known combination of
zonally symmetrical annular mode with zonal asymmetries in the form of a
wave-3 \citep{fogt2012}. The regression using the Asymmetric and
Symmetric SAM indices successfully disentangle both structures. The
Asymmetric SAM index gives rise to a cleaner zonal wave (panel b.2) and
the Symmetric SAM index is associated with an annular mode, almost
devoid of zonal asymmetries (panel c.2). The wave-3 pattern observed in
panel b.2 is rotated by half a wavelength from the average position of
the mean wave-3 pattern associated with \citet{raphael2004}'s ZW3 index,
whose reference locations are marked with points in the figure. Thus,
the tropospheric Asymmetric SAM index represents a zonal displacement in
the position of the climatological wave-3 pattern.

\begin{figure*}
\includegraphics{wave-amplitude-1} \caption{Amplitude (meters) of zonal waves of the geopotential height regression patterns in Figure \ref{fig:2d-regr} for zonal waves with wave-number 0, 1, 2, and 3, where wave-number 0 represents the amplitude of the zonal mean.}\label{fig:wave-amplitude}
\end{figure*}

The amplitude of the first zonal wave numbers at each latitude at 50 hPa
and 700 hPa is shown in Figure \ref{fig:wave-amplitude}, where wave
number zero represents the amplitude of the zonal mean. Zonal wave
amplitudes of the spatial pattern described by the Full SAM index
(column a) are dominated by the zonal mean (wave-number 0) at both
levels. However, zonal waves are important, particularly North of
50\degree S, with wave-number 1 clearly dominating at 50 hPa (panel a.1)
and a more equal mix of waves at 700 hPa (panel a.2). Column b shows
that the Asymmetric SAM is overwhelmingly dominated by wave 1 in the
stratosphere (panel b.1), while in the troposphere it is composed of
zonal waves 3 to 1 in decreasing level of importance (panel b.2) with
negligible amplitude of the zonal mean. The Symmetric SAM, on the other
hand, it's almost entirely composed of zonal mean at both levels (column
c), with little to now contribution from zonal waves with wave-numbers 1
to 3.

We can see that the amplitude and latitudinal distribution zonal waves
in the Asymmetric SAM on one hand, and the zonal mean in the Symmetric
SAM on the other correspond almost exactly to the amplitude and
latitudinal distribution in the Full SAM. This confirms the correct
decomposition of the SAM in its symmetric and asymmetric components.

Looking at panel b.2 from Figure \ref{fig:2d-regr}, it becomes apparent
that zonal waves 1 and 2 modulate the amplitude of zonal wave 3, which
--as mentioned before-- is larger in the Western Hemisphere than in the
Eastern Hemisphere.

\begin{figure}
\includegraphics{vertical-regression-1} \caption[Regression between monthly geopotential height anomalies (meters) averaged betweeen 65\degree and 40\degree S and the Asymmetric SAM index (extracted from multiple regression including the Symmetric SAM)]{Regression between monthly geopotential height anomalies (meters) averaged betweeen 65\degree and 40\degree S and the Asymmetric SAM index (extracted from multiple regression including the Symmetric SAM). (a) With the Asymmetric SAM in 50 hPa and (b) in 700 hPa for the 1979–2018 period.}\label{fig:vertical-regression}
\end{figure}

To analyse the vertical structure of the geopotential height anomalies
associated with the asymmetric SAM index, we show a vertical cross
section of regressions of mean geopotential height between 65\degree S
and 40\degree S for the 50 hPa Asymmetric SAM index (panel a) and for
the 700 hPa Asymmetric SAM index (panel b) (Figure
\ref{fig:vertical-regression}). The geopotential height anomalies
associated with the stratospheric Asymmetric SAM (panel a) are clearly
constrained to the stratosphere, which underscores the uncoupling
between the stratospheric and tropospheric Asymmetric SAM. The vertical
structure of this signal tilts about 60\degree to the West between 100
hPa and 1 hPa, suggesting baroclinic processes. Interestingly, the
signal in the stratosphere maximises near 10 hPa despite using the 50
hPa index for the regression.

The tropospheric Asymmetric SAM (panel b) has significant signals that
extend upwards to the uppermost levels of the reanalysis. In the
troposphere, the wave-3 structure is equivalent barotropic with maximum
amplitude at roughly 250 hPa. The anomalies are much more intense in the
Western hemisphere, where they extent into the stratosphere. In the
Eastern hemisphere the wave-3 signal is weaker and confined to the
troposphere while negative anomalies dominate in the stratosphere. So,
while the tropospheric Asymmetric SAM index is associated with
stratospheric geopotential anomalies, these do not project strongly onto
the stratospheric Asymmetric SAM. The structures shown in panels a and b
in Figure \ref{fig:vertical-regression} are robust to the choice of
index level. For any stratospheric (above 100 hPa) index, the resulting
anomalies are very similar to the wave-1 structure with maximum near 10
hPa in panel a. Conversely, for any tropospheric (below 100 hPa) index,
the result is very similar to panel b. The patterns mainly change in
amplitude (not shown).

The wave-3 pattern from Figure \ref{fig:2d-regr} panel b.2 is very
similar to the Pacific-South American Pattern \citep{mo1987, kidson1988}
which is a teleconnection pattern associated with the ENSO
\citep{karoly1989}. Indeed, \citet{fogt2011} showed that there is a
significant relationship between the SAM and the ENSO. The correlation
between the Full SAM and the ENSO as measured by the Oceanic Niño Index
\citep{bamston1997} (ONI) is -0.16 (p-value =
\ensuremath{2.8\times 10^{-4}}). Consistent with \citet{fan2007}, we
show that this relationship is captured entirely by the Asymmetric SAM,
as this index has a partial correlation of -0.26 (p-value =
\ensuremath{6.3\times 10^{-9}}) with the ONI controlling for the effect
of the Symmetric SAM, whereas the Symmetric SAM's partial correlation
with the ONI is essentially null (0.019; p-value = 0.67). We performed
the same analysis using the Multivariate ENSO Index \citep{wolter2011}
and the Southern Oscillation Index \citep{ropelewski1987} to conclude
that these results do not depend on the ENSO index used.

\subsection{Impacts}

\label{sec:impacts}

\begin{figure*}
\includegraphics{regr-air-season-1} \caption[Regression of seasonal mean surface temperature (Kelvin) with Asymmetric SAM and Symmetric SAM for the 1979–2018 period]{Regression of seasonal mean surface temperature (Kelvin) with Asymmetric SAM and Symmetric SAM for the 1979–2018 period. Black contours indicate areas with p-value smaller than 0.05 controlling for False Detection Rate. Gray areas in Antarctica are areas with have more than 15\% of missing data.}\label{fig:regr-air-season}
\end{figure*}

To see if there are different surface impacts associated with the
asymmetric and symmetric SAM circulation we regress surface temperature
and precipitation onto each of the three SAM indices at 700 hPa. As
shown in previous sections, the three indices are highly coherent in the
troposphere, so we select this level to represent the tropospheric
circulation for consistency with previous studies.

Figure \ref{fig:regr-air-season} shows regression coefficients of each
index at 700 hPa with surface temperature for each trimester. In Summer,
positive values of the Full SAM index (panel a.1) are associated with
negative temperature anomalies near Antarctica which are surrounded by a
ring of positive anomalies. The ring is not zonally symmetric, as there
are four clear local maximums around 30\degree W, 120\degree W,
150\degree E and 90\degree E. In the tropics, there are negative
anomalies in the equatorial Pacific, consistent with the negative
correlation between SAM and ENSO. Panels b.1 and c.1 show temperature
anomalies associated with positive values of the Asymmetric and
Symmetric SAM, respectively. Both the local maximums in the ring and the
anomalies in the Pacific regions are present mostly on the Asymmetric
SAM regression map, while temperature patterns linked to positive
Symmetric SAM show a more zonally consistent ring and less relation to
the tropics. Noticeable, temperature anomalies in the Indian ocean,
South Africa and Australia are strongly related to Asymmetric SAM. This
signal is not present in the regression pattern with the Full SAM.
Spring (row 4) features similar patterns but of smaller magnitude, with
less regions where regressed anomalies have statistical significance.

In Autumn and Winter (rows 2 and 3) the positive ring is only present
through its local maximums in the regression with the Full SAM. There
are also negative anomalies in Southern Australia, and positive
anomalies over New Zealand and Southern South America. These patterns
are not significant in the sense that there are no areas with p-values
below 0.05 when controlling for FDR following \citet{wilks2016}.
However, repeating this analysis with 2-meter temperature from ERA5
resulted in similar patterns that were statistically significant (not
shown). Moreover, similar features were observed in station measurements
by \citet{jones2019}, although using data from 1957 to 2016.

The pattern of negative anomalies in the pole surrounded by positive
anomalies roughly seen in all seasons --although with varying intensity
and small-scale details-- translates to enhanced meridional temperature
gradient maximised in the zero line, which is consistent with the
intensification and poleward migration of the westerlies commonly linked
to the SAM through thermal wind balance. It's then not surprising to see
it more clearly in association with the Symmetric SAM (at least in
Summer and Spring).

Figure \ref{fig:regr-air-season} column b can be partially compared with
Figure 11 from \citet{fogt2012}. Although they used station data from
1958 to 2001, main features are reproduced here, such as the strong
signal in New Zealand and Australia in Summer and Spring.

Regression of the SAM indices with seasonal mean precipitation and 700
hPa geopotential height are shown in Figures \ref{fig:pp-regr-oceania}
and \ref{fig:pp-regr-america} for Australasia and South America
respectively. South Africa is not shown because no significant signal
was detected there.

\begin{figure*}
\includegraphics{pp-regr-oceania-1} \caption[Regression of (row 1) annual and (rows 2 to 5) seasonal mean precipitation anomalies (mm per day, shading) and 700 hPa geopotential height (thin lines, positive values as solid lines and negative values as dashed lines) with (column a) Full SAM, (column (b) Asymmetric SAM and (column c) Symmetric SAM for the 1979–2018 period]{Regression of (row 1) annual and (rows 2 to 5) seasonal mean precipitation anomalies (mm per day, shading) and 700 hPa geopotential height (thin lines, positive values as solid lines and negative values as dashed lines) with (column a) Full SAM, (column (b) Asymmetric SAM and (column c) Symmetric SAM for the 1979–2018 period. Thin lines are the Black contours indicate areas with p-value smaller than 0.05 controlling for False Detection Rate.}\label{fig:pp-regr-oceania}
\end{figure*}

In Australia, the annual regression shows that the Full SAM index is
positively associated with precipitation in the Southeastern region
(Figure \ref{fig:pp-regr-oceania} panel a.1), which reproduces the
results from \citet{gillett2006}. The separation between Asymmetric and
Symmetric SAM suggest that this positive anomaly is explained by the
Symmetric SAM only in the East coast (panel c.1). Geopotential anomalies
associated with this index (black contours) are indicative of easterly
flow from the Tasman Sea, which could explain the positive anomalies in
precipitation as found by \citet{hendon2007}. The Asymmetric SAM appears
related to increased precipitation in the West coast of Southeastern
Australia (panel b.2), which could similarly be explained by the
anomalous westerly circulation transporting moist air to the continent
from the Indian Ocean.

The seasonal-level regressions show statistically significant anomalies
only in Spring, when positive Full SAM is associated with positive
precipitation anomalies in Eastern Australia (panel a.5). In this
trimester the Symmetric SAM seems to be associated with precipitation in
a relatively reduced area of the East Coast (panel c.5) while the
positive precipitation anomalies related with positive Asymmetric SAM
affect all Eastern Australia (panel b.5).

In Summer, positive Full SAM index is associated with with positive
precipitation anomalies in Western and Eastern Australia, particularly
in the North East (panel a.2). The Eastern part being dominated by the
relationship with the Symmetric SAM and the Western, by the Asymmetric
SAM. In Autumn, the regression with Full SAM shows positive values in
the North, similar to Summer, and a broad area of positive values in the
North-East to South-West direction. This structure seems to be
associated with the Symmetric SAM, while the Northern positive values
are associated with the Asymmetric SAM. In Winter we see the same NE to
SW aligned anomaly (although with much reduced amplitude) that is also
present only in relation with the Symmetric SAM. None of these
regression coefficients are statistically significant at the 95\% level.
The Spring signal is broadly consistent with \citet{hendon2007}, but
whereas \citet{hendon2007} also detected a strong signal in Summer,
panel a.2 shows no statistically significant association (although the
coefficients have the consistent sign).

\begin{figure*}
\includegraphics{pp-regr-america-1} \caption{Same as Figure \ref{fig:pp-regr-oceania} but for South America.}\label{fig:pp-regr-america}
\end{figure*}

In South America (Figure \ref{fig:pp-regr-america}), the annual-level
regression shows that positive SAM is associated with statistically
significant precipitation decrease in Southeastern South America (SESA)
and Southern Chile and non-significant increase in South Brazil, near
the South Atlantic Convergence Zone (SACZ) (panel a.1). Panels b.1 and
c.1 show a remarkably clean separation between the Asymmetric SAM
--associated with the Southeastern South American and Southern Brazilian
signals-- and the Symmetric SAM --associated with the signal in Southern
Chile.

Except Winter, seasonal-level regressions mirror this same pattern. Even
if not statistically significant, they all show negative values in
Southeastern South America and Southern Chile along with positive values
in Southern Brazil in relation with the Full SAM. The separation of
these features between the Asymmetric SAM and Symmetric SAM regression
maps is also rather consistent.

The anomalous circulation at 700 hPa associated with the Symmetric SAM
(panel c.1) indicate anomalous Easterly flow over Southern Chile. This
leads to reduced influx of moist air from the Pacific Ocean which, is
the main source of precipitable water in that region
\citep{garreaud2007}. On the other hand, the anomalous circulation
associated with positive values of Asymmetric SAM (panel b.1) in the
Atlantic is anticyclonic in the South and cyclonic in the North. This
creates anomalous South-Easterly flow over Southeastern South America,
which inhibits the flow of the Low Level Jet to the region
\citep[\citet{zamboni2010}]{silvestri2009}. This same pattern was found
to be associated with increased precipitation in Southern Brazil during
South Atlantic Convergence Zone events \citep{rosso2018}. There is a
small area of increased precipitation with SAM near central Argentina
which is also present in the station-based analysis by
\citet{gillett2006} and that is explained by the Asymmetric SAM.

\subsection{Conclusions}

In this study we characterise the temporal and spatial variability of
the zonally symmetric and asymmetric structure of the SAM. By projecting
monthly geopotential fields at each level with the corresponding
asymmetric and symmetric pattern, we created two indices for
representing the zonally asymmetric and symmetric contributions of the
SAM respectively.

As expected, the Asymmetric SAM index correlates strongly with the
Symmetric SAM index. In the troposphere, this correlation is maximum at
zero lag, while in the stratosphere is maximised with the Asymmetric SAM
leading the Symmetric SAM by one month. Since most indices of the SAM
are calculated using surface or near-surface conditions, this result
would suggest that they might not be sensitive to the most dramatic
changes in SAM variability.

The two-year periodicity we found in the stratospheric Symmetric SAM
might point to a link between the SAM and the Quasi Biennial
Oscillation. There is evidence of influence between the QBO and the
Northern Annular Mode \citep[e.g.][\citet{watson2014},
\citet{zhang2020}]{holton1980}, so it's not unlikely that the SAM would
be similarly connected. However establishing this link would require
further research.

We observe a positive trend towards positive SAM in Summer and Autumn,
As was documented by previous studies, such as \citet{fogt2020} (and
references therein) for surface levels. We show that these trends are
maximised at the 100 hPa level and are explained by the zonally
symmetric component. We also find a statistically significant positive
trend in the Symmetric component of the SAM in the stratosphere that is
not apparent in the Full SAM index. In contrast to \citet{fogt2012} we
find some evidence of the SAM becoming more zonally asymmetric, as there
is a slight positive trend in the variance explained by the as the
Asymmetric SAM explains an increasingly proportion of the total
variance.

In the troposphere, the spatial patterns of geopotential associated with
the Symmetric SAM are much closer to being truly annular than the
patterns associated with the Full SAM index. The Asymmetric SAM, on the
other hand, describes a wave-3 pattern with maximum amplitude in the
Pacific region and whose phase is rotated a quarter wavelength from the
mean zonal wave 3 described by \citet{raphael2004}'s index. This pattern
extends in the troposphere but its maximum is located at 250 hPa, which
also could suggest that surface-based indices are not optimum for
capturing this variability.

This wave-3 pattern is similar to the Pacific-South American Pattern,
which is a teleconnection pattern linked to ENSO variability. We found
that the significant correlation that exists between the Full SAM index
and the Oceanic Niño Index is captured entirely by the Asymmetric SAM
index. This suggests that ENSO is linked to SAM exclusively through the
variability in the latter's asymmetric component and thus, the
Asymmetric SAM index could be a useful measure to further study that
relationship.

Temperature anomalies associated with the Full SAM broadly show a
pattern of negative anomalies at polar latitudes surrounded by positive
anomalies, but with many deviations from symmetry. The Asymmetric SAM
index explains a big portion of these deviations. In particular, the
positive phase of the Asymmetric SAM is associated with colder
temperatures over Southern Brazil, South Africa and Southern Australia,
as well as the negative anomalies in the equatorial Pacific consistent
with the ENSO-SAM relationship. These negative anomalies are
particularly clear in the DJF and SON trimesters, which include the
months in which the ENSO teleconnection is more active
\citep{cazes-boezio2003, fogt2011, cai2020a}.

In Australia the Full SAM is associated with positive precipitation
anomalies in South East and this is explained by the Symmetric SAM.
However, the Asymmetric SAM is associated with a small area of positive
precipitation anomalies in the Eastern Coast of West Australia, maybe
due to advection of moist air from the Indian Ocean.In South America,
precipitation anomalies associated with the Full SAM are negative both
in Southern Chile and Southeastern South America, and positive in
Southern Brazil. This features are cleanly separated between the
Asymmetric and Symmetric components. The Symmetric SAM explains the
negative anomalies in Southern Chile and the Asymmetric SAM, the
negative-positive dipole between Southeastern South America and Southern
Brazil. Individual seasons mostly follow this pattern.

\citet{silvestri2009} suggests that precipitation impacts linked to the
SAM changed rather dramatically before and after 1980. In particular,
the negative relationship with precipitation in South America was absent
in some areas and switched sign in others in the earlier period. The
correlation between ENSO and SAM is similarly non-stationary, also
changing sign before the 1980s \citep{fogt2006, clem2013}. Seeing as
both the ENSO-SAM relationship and most of the precipitation impacts in
South America are captured by the Asymmetric SAM, the results presented
here are most likely period-dependent.

By succesfully separating the zonally symmetric and zonally symmetric
SAM signals, we show that the asymmetric component of the SAM has its
unique variability, trends and impacts. This is particularly important
in the context of a changing climate, as the impact on the SAM of ozone
recovery is modeled as highly zonally symmetric, while the impact of
increased concentration of greenhose gases has also a zonally asymmetric
component \citep{arblaster2006, simpkins2012}.

\acknowledgments

NOAA Global Surface Temperature (NOAAGlobalTemp) data provided by the
NOAA/OAR/ESRL PSL, Boulder, Colorado, USA, from their Web site at
https://psl.noaa.gov/

The research was supported by UBACyT20020170100428BA and the CLIMAX
Project funded by Belmont Forum/ANR-15-JCL/-0002-01. Elio Campitelli was
supported by a PhD grant from CONICET, Argentina.

\bibliography{AsymSAM,packages}

\newpage

\appendix

\appendixtitle{Extra figures}

\begin{figure}
\includegraphics{A1-1} \appendcaption{A1}{Lag-correlation between Asymmetric SAM and Symmetric SAM index at each level. Negative lags imply Symmetric SAM leading Asymmetric SAM and vice versa.}\label{fig:A1}
\end{figure}

\begin{figure}
\includegraphics{A3-1} \appendcaption{A3}{Fourier spectrum of each timeseries computed as Fourier transform smoothed with modified Daniell smoothers with widths 3 and 5. The shading indicates de 95\% confidence area derived by fitting an autorregressive model and computing the spectrum for 5000 simulated samples from the fitted autoregressive model (95\% of the simulated sampels had an amplitude equal or lower). The light line indicates the theoretical expected amplitude from the autorregressive model.}\label{fig:A3}
\end{figure}

\begin{figure}
\includegraphics{A9-1} \caption[700 hPa Geopotetnial height zonal anomalies (meters) of composites of positive and negative SAM months selected using $\pm1$ standard deviation as threshhold]{700 hPa Geopotetnial height zonal anomalies (meters) of composites of positive and negative SAM months selected using $\pm1$ standard deviation as threshhold. Numbers in the column headers are pattern correlation between SAM+ and SAM- composites and number of monthly fields used to construct the composites.}\label{fig:A9}
\end{figure}

\begin{figure}
\includegraphics{A10-1} \caption[Regression of 700 hPa geopotential height zonal anomalies (meters) onto the standarised timeseries of the leading EOF computed for each season independently]{Regression of 700 hPa geopotential height zonal anomalies (meters) onto the standarised timeseries of the leading EOF computed for each season independently.}\label{fig:A10}
\end{figure}

\begin{figure}
\includegraphics{A11-1} \caption[Regression of 700 hPa geopotential height zonal anomalies (meters) onto the standarised timeseries of the leading EOF computed for the periods 1979 to 1998 and 1999 to 2018]{Regression of 700 hPa geopotential height zonal anomalies (meters) onto the standarised timeseries of the leading EOF computed for the periods 1979 to 1998 and 1999 to 2018. Pattern correlation between both fields is 0.76.}\label{fig:A11}
\end{figure}

\end{document}